---
title: "02_Setbp1_Community_Identification"
author: "Jordan Whitlock"
date: '2023-06-28'
output: html_document
---

Goal: Determine what Community Number TF Setbp1 is contained in for every single cell type in both tissues. This is needed for JI comparisons between Sebtp1 communities in following script.

```{r}
set.seed(2178)
library(here)
library(tidyverse)
library(styler)
library(dplyr)
ptm <- proc.time()
```

```{r}
# load in gene set for Setbp1
setbp1_genes <- read.csv(here("results/seurat/setbp1_targets.csv"))
setbp1_genes <- setbp1_genes[, -1] %>% as.vector()
setbp1_genes <- append(setbp1_genes, "Setbp1")
```

```{r cortex-diff-mod-comm-members-of-setbp1}
pattern <- "_cortex.RData_ALPACA_scores.txt"
directory <- here("results/alpaca/main")
files <- list.files(directory, pattern = paste0(".*", pattern), full.names = TRUE)
toevaluate <- sub("^(.*?)_ALPACA.*$", "\\1", basename(files))



scores_list <- list()

for (t in 1:length(toevaluate)) {
  name <- sub("^(.*?)_cortex\\.RData.*$", "\\1", toevaluate[t])
  # Load ALPACA scores and final communities
  readscores <- paste(here("results/alpaca/main/"), toevaluate[t], "_ALPACA_scores.txt", sep = "")
  readcomms <- paste(here("results/alpaca/main/"), toevaluate[t], "_ALPACA_final_memb.txt", sep = "")
  scores <- read.delim(readscores, header = F)
  comms <- read.delim(readcomms, header = F)
  tosel <- intersect(scores[, 1], comms[, 1])
  scores <- scores[which(scores[, 1] %in% tosel), ]
  comms <- comms[which(comms[, 1] %in% tosel), ]
  scores <- scores[order(scores[, 1]), ]
  comms <- comms[order(comms[, 1]), ]
  scores[, 1] <- as.character(scores[, 1])
  comms[, 1] <- as.character(comms[, 1])
  all(scores[, 1] == comms[, 1])
  scores <- cbind(scores, comms[, 2])
  row.names(scores) <- scores[, 1]
  scores <- scores[, -1]
  colnames(scores) <- c("score", "com")
  scores_list[[name]] <- scores
}

# creating column for gene/TF name:
for (i in seq_along(scores_list)) {
  # Get the current dataframe
  df <- scores_list[[i]]

  # Move row names to a column called "name"
  df$name <- rownames(df)

  # Split the column into two separate columns
  df <- df %>%
    separate(name, into = c("name", "entity"), sep = "_", remove = FALSE)

  # Replace "_A" with "TF" and "_B" with "gene" in the second column
  df$entity <- ifelse(df$entity == "A", "TF", "gene")

  # Reset row names to NULL
  rownames(df) <- NULL

  # Update the dataframe in the list
  scores_list[[i]] <- df
}
```

```{r cortex-diff-mod-comm-members-of-setbp1-targets}
# Subset each dataframe for Setbp1 gene set in the list and annotate with the list name
subset_dfs <- lapply(names(scores_list), function(list_name) {
  df <- scores_list[[list_name]]
  subset_df <- df[df$name %in% setbp1_genes, ]
  subset_df$cell_type <- list_name
  return(subset_df)
})

# Bind all subset dataframes into a single dataframe
setbp1_communities <- bind_rows(subset_dfs)

# save
write.csv(setbp1_communities, file = here("results/alpaca/cortex_communities_setbp1_geneset.csv"), row.names = FALSE)

# filter for Setbp1 to determine its community number for each cell type for Jaccard Index analysis
# Filter rows where "name" is "Setbp1"
filtered_df <- setbp1_communities %>%
  filter(name == "Setbp1" & entity == "TF")

# Group by "cell_type" and output values in "com" column
output <- filtered_df %>%
  group_by(cell_type) %>%
  summarize(com_value = unique(com))

# Print the output
print(output) # Setbp1 is in community 1 for all cell types in cortex
```

```{r kidney-diff-mod-comm-members-of-setbp1}
pattern <- "_kidney.RData_ALPACA_scores.txt"
directory <- here("results/alpaca/main")
files <- list.files(directory, pattern = paste0(".*", pattern), full.names = TRUE)
toevaluate <- sub("^(.*?)_ALPACA.*$", "\\1", basename(files))



scores_list <- list()

for (t in 1:length(toevaluate)) {
  name <- sub("^(.*?)_kidney\\.RData.*$", "\\1", toevaluate[t])
  # Load ALPACA scores and final communities
  readscores <- paste(here("results/alpaca/main/"), toevaluate[t], "_ALPACA_scores.txt", sep = "")
  readcomms <- paste(here("results/alpaca/main/"), toevaluate[t], "_ALPACA_final_memb.txt", sep = "")
  scores <- read.delim(readscores, header = F)
  comms <- read.delim(readcomms, header = F)
  tosel <- intersect(scores[, 1], comms[, 1])
  scores <- scores[which(scores[, 1] %in% tosel), ]
  comms <- comms[which(comms[, 1] %in% tosel), ]
  scores <- scores[order(scores[, 1]), ]
  comms <- comms[order(comms[, 1]), ]
  scores[, 1] <- as.character(scores[, 1])
  comms[, 1] <- as.character(comms[, 1])
  all(scores[, 1] == comms[, 1])
  scores <- cbind(scores, comms[, 2])
  row.names(scores) <- scores[, 1]
  scores <- scores[, -1]
  colnames(scores) <- c("score", "com")
  scores_list[[name]] <- scores
}

# creating column for gene/TF name:
for (i in seq_along(scores_list)) {
  # Get the current dataframe
  df <- scores_list[[i]]

  # Move row names to a column called "name"
  df$name <- rownames(df)

  # Split the column into two separate columns
  df <- df %>%
    separate(name, into = c("name", "entity"), sep = "_", remove = FALSE)

  # Replace "_A" with "TF" and "_B" with "gene" in the second column
  df$entity <- ifelse(df$entity == "A", "TF", "gene")

  # Reset row names to NULL
  rownames(df) <- NULL

  # Update the dataframe in the list
  scores_list[[i]] <- df
}
```

```{r kidney-diff-mod-comm-members-of-setbp1-targets}
# Subset each dataframe for Setbp1 gene set in the list and annotate with the list name
subset_dfs <- lapply(names(scores_list), function(list_name) {
  df <- scores_list[[list_name]]
  subset_df <- df[df$name %in% setbp1_genes, ]
  subset_df$cell_type <- list_name
  return(subset_df)
})

# Bind all subset dataframes into a single dataframe
setbp1_communities <- bind_rows(subset_dfs)

# save
write.csv(setbp1_communities, file = here("results/alpaca/kidney_communities_setbp1_geneset.csv"), row.names = FALSE)

# filter for Setbp1 to determine its community number for each cell type for Jaccard Index analysis
# Filter rows where "name" is "Setbp1"
filtered_df <- setbp1_communities %>%
  filter(name == "Setbp1" & entity == "TF")

# Group by "cell_type" and output values in "com" column
output <- filtered_df %>%
  group_by(cell_type) %>%
  summarize(com_value = unique(com))

# Print the output
print(output) # Setbp1 is in community 1 for all cell types in kidney
```

```{r processing-time}
fptm <- proc.time() - ptm
fptm[3] / 60
```
  elapsed 
0.1966667 


```{r style-file}
# run style
style_file(here("src/network_scripts/community_detection/02_Setbp1_Community_Identification.Rmd")) # commented out after being run once
# lintr was run as well
```

```{r session-info}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.6 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] styler_1.9.0    lubridate_1.9.2 forcats_1.0.0   stringr_1.5.0   dplyr_1.1.0     purrr_1.0.1     readr_2.1.4     tidyr_1.3.0     tibble_3.1.8    ggplot2_3.4.1  
[11] tidyverse_2.0.0 here_1.0.1     

loaded via a namespace (and not attached):
  [1] Seurat_4.3.0           Rtsne_0.16             colorspace_2.1-0       deldir_1.0-6           ellipsis_0.3.2         ggridges_0.5.4        
  [7] rprojroot_2.0.3        spatstat.data_3.0-0    rstudioapi_0.14        leiden_0.4.3           listenv_0.9.0          ggrepel_0.9.3         
 [13] fansi_1.0.4            R.methodsS3_1.8.2      codetools_0.2-18       splines_4.1.3          knitr_1.42             polyclip_1.10-4       
 [19] jsonlite_1.8.4         ica_1.0-3              cluster_2.1.2          R.oo_1.25.0            png_0.1-8              uwot_0.1.14           
 [25] spatstat.sparse_3.0-0  shiny_1.7.4            sctransform_0.3.5      compiler_4.1.3         httr_1.4.5             SeuratObject_4.1.3    
 [31] Matrix_1.5-3           fastmap_1.1.1          lazyeval_0.2.2         cli_3.6.0              later_1.3.0            htmltools_0.5.4       
 [37] tools_4.1.3            igraph_1.4.1           gtable_0.3.1           glue_1.6.2             RANN_2.6.1             reshape2_1.4.4        
 [43] Rcpp_1.0.10            scattermore_0.8        vctrs_0.5.2            nlme_3.1-155           spatstat.explore_3.0-6 progressr_0.13.0      
 [49] lmtest_0.9-40          spatstat.random_3.1-3  xfun_0.37              globals_0.16.2         timechange_0.2.0       mime_0.12             
 [55] miniUI_0.1.1.1         lifecycle_1.0.3        irlba_2.3.5.1          goftest_1.2-3          future_1.31.0          MASS_7.3-55           
 [61] zoo_1.8-11             scales_1.2.1           spatstat.utils_3.0-1   hms_1.1.2              promises_1.2.0.1       parallel_4.1.3        
 [67] RColorBrewer_1.1-3     yaml_2.3.7             reticulate_1.28        pbapply_1.7-0          gridExtra_2.3          stringi_1.7.12        
 [73] rlang_1.0.6            pkgconfig_2.0.3        matrixStats_0.63.0     evaluate_0.20          lattice_0.20-45        tensor_1.5            
 [79] ROCR_1.0-11            patchwork_1.1.2        htmlwidgets_1.6.1      cowplot_1.1.1          tidyselect_1.2.0       parallelly_1.34.0     
 [85] RcppAnnoy_0.0.20       plyr_1.8.8             magrittr_2.0.3         R6_2.5.1               generics_0.1.3         pillar_1.8.1          
 [91] withr_2.5.0            fitdistrplus_1.1-8     abind_1.4-5            survival_3.3-1         sp_1.6-0               future.apply_1.10.0   
 [97] crayon_1.5.2           KernSmooth_2.23-20     utf8_1.2.3             spatstat.geom_3.0-6    plotly_4.10.1          tzdb_0.3.0            
[103] rmarkdown_2.20         grid_4.1.3             data.table_1.14.8      digest_0.6.31          R.cache_0.16.0         xtable_1.8-4          
[109] httpuv_1.6.9           R.utils_2.12.2         munsell_0.5.0          viridisLite_0.4.1     
