---
title: "README"
author: "Jordan Whitlock"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Network Construction and Analysis

## Purpose:
All scripts in this directory are used to build necessary inputs required for network construction as well as downstream analyses.

### Reproducibility:

**Network Construction**
* PANDA jobs were run using UAB Cheaha Supercomputer and it's SLURM scheduler. Jobs were run as an array, within a [Docker](https://hub.docker.com/repository/docker/jordanwhitlock/jw_diseasenetworks/general) container converted Singularity in order to execute.

* Expression data was obtained from GTEx using Recount3 (accessed: 220806-220807) and includes all tissue types except blood and study_na
* Human PPI and TF Motif input construction is detailed [here](https://github.com/lasseignelab/230323_JW_DiseaseNetworks/tree/main/src/panda_input_construction).

### Scripts:

#### PANDA Expression Inputs and Network Construction:
This directory contains all scripts needed to construct [PANDA](https://netzoo.github.io/zooanimals/panda/) regulatory networks. Prior to building PANDA networks, the user must first construct the expression inputs needed and then build the *.txt*, which is a list of expression input file paths needed for the array job.

```{r echo = FALSE}
fs::dir_tree()
```