---
title: "01_tissue_specific_edge_identification"
author: "Jordan Whitlock"
date: '2023-08-01'
output: html_document
---

```{r}
set.seed(2178)
library(here)
library(tidyr)
library(stringr)
library(lintr)
library(ggplot2)
library(styler)
library(dplyr)
library(reshape2)
proc.time()
```

```{r}
# affected tissues
affected <- c(
  "BLOOD",
  "BRAIN",
  "HEART",
  "KIDNEY",
  "BLADDER",
  "LUNG",
  "MUSCLE",
  "SMALL_INTESTINE",
  "STOMACH",
  "ESOPHAGUS",
  "BONE_MARROW"
)
```

```{r loading-geneset}
# load in gene set for SETBP1 (converted by LW)
setbp1_genes <- read.csv(here("data/SETBP1_Targets/setbp1_targets_geneconversions.csv"))
setbp1_genes <- setbp1_genes$name %>% as.vector()
```


# generating edges from the regNet
```{r}
files <- list.files(here("results/PANDA"), full.names = TRUE)

names <- sub("expression_PANDA.Rdata", "", basename(files)) %>%
  sub("gtex_", "", .) %>%
  sub("_PANDA.Rdata", "", .)

regNet_edges <- list()

for (i in files) {
  # load in the pandaResults
  load(i)
  # extracting-networks
  regNet <- pandaResults@regNet
  regNet <- reshape2::melt(regNet, varnames = c("TF", "gene"), value.name = "edge_weight")

  # filter for TF SETBP1
  regNet <- regNet[regNet$TF == "SETBP1", ]

  # creating edges
  regNet$edges <- paste(regNet$TF, regNet$gene, sep = "_")
  regNet <- subset(regNet, select = -c(TF, gene))

  # store in list
  regNet_edges[[i]] <- regNet

  # print
  print(paste0(i, "melted and edges created"))
}

names(regNet_edges) <- names

save(regNet_edges, file = here("results/tissue_specific_edges/edges_list_object.Rdata"))
```

# combine all regNets for all tissues 
```{r}
merged_df <- regNet_edges[[1]]
for (i in 2:length(regNet_edges)) {
  merged_df <- left_join(merged_df, regNet_edges[[i]], by = "edges")
}

# reorder the columns
merged_df <- merged_df %>%
  select(edges, edge_weight, everything())

# rename the columns
names(merged_df)[-1] <- names(regNet_edges)

# save the object
save(merged_df, file = here("results/tissue_specific_edges/merged_edges_object.Rdata"))
```

We wanted to identify tissue-specific edges as previously defined in [Sonawane et al Cell Reports 2017](https://www.cell.com/cell-reports/fulltext/S2211-1247(17)31418-3#secsectitle0080); formulas 1, 2 & 3 to calculate tissue-specific-edges
```{r}
# ln(e^x + 1) transform edge weights to address negatives
merged_df <- merged_df %>% mutate(across(-1, ~ log(exp(.) + 1)))

save(merged_df, file = here("results/tissue_specific_edges/merged_edges_transformed.Rdata"))

# make df for storing median and iqr
med_iqr_df <- data.frame(edges = merged_df$edges)

# calculate median and IQR for each edge (row) across all tissues
med_iqr_df$median_all <- apply(merged_df[, -1], 1, median) # median for each row
med_iqr_df$IQR_all <- apply(merged_df[, -1], 1, IQR) # IQR for each row

save(med_iqr_df, file = here("results/tissue_specific_edges/med_iqr_df.Rdata"))

# compare edge weights across tissues using Sonawane et al Equation 2 to calculate specificity scores
merged_df[, -1] <- t(apply(merged_df[, -1], 1, function(x) (x - med_iqr_df$median_all) / med_iqr_df$IQR_all)) # transpose because apply returns rows as tissues and columns as edges

# calculate multiplicity
merged_df$multiplicity <- apply(merged_df[-1], 1, function(x) sum(x > 2))
hist(merged_df$multiplicity) # visualize

multiplicity <- merged_df[, c(1, 33)]

save(merged_df, file = here("results/tissue_specific_edges/specificity_multiplicity.Rdata"))

multiplicity <- multiplicity %>%
  separate(edges, into = c("TF", "gene"), sep = "_", remove = FALSE) # split edges

save(multiplicity, file = here("results/tissue_specific_edges/multiplicity.Rdata"))
```

# define tissue specific edges as those with a specificity scores > 2
```{r}
merged_df <- merged_df[, -33]
merged_df[, -1] <- lapply(merged_df[, -1], function(x) ifelse(x > 2, 1, 0))

# melt tissue-specific edges and grab only those that are tissue-specific
tissue_specific <- melt(merged_df, id.vars = "edges",
                        variable.name = "Tissue",
                        value.name = "Tissue Specific")
tissue_specific <- tissue_specific[tissue_specific$`Tissue Specific` == 1, ]

# split edges
tissue_specific <- tissue_specific %>%
  separate(edges, into = c("TF", "gene"), sep = "_", remove = FALSE)

save(tissue_specific, file = here("results/ts_edges/tissue_specific_edges.Rdata"))
```

## Cross reference tissue-specific edges with the CONDOR communitites for SETBP1
```{r}
load(here("results/condor/condor_list.Rdata"))
setbp1_comm_number <- read.csv(here("results/condor/communitynumber_setbp1_TF.csv"))
```

```{r comm-members-tse-edges}
# grab nodes interacting with SETBP1 from tse
tse_genes <- as.vector(unique(tissue_specific$gene))

# Subset each dataframe for tse genes in the list and annotate with the list name
subset_dfs <- lapply(names(condor_list), function(list_name) {
  df <- condor_list[[list_name]]
  subset_df <- df[df$name %in% tse_genes, ]
  subset_df$tissue_type <- list_name
  return(subset_df)
})

# Bind all subset dataframes into a single dataframe
tse_edges <- bind_rows(subset_dfs)

# filter communities for communities known to contain SETBP1 for each tissue
setbp1_comms <- paste(setbp1_comm_number$tissue_type, setbp1_comm_number$com_value, sep = "_")# generate SETBP1 community vector label
tse_edges$annotation <- paste(tse_edges$tissue_type, tse_edges$condor_memb, sep = "_")# generate annotation to filter tse edges for setbp1_comms with
data <- tse_edges[tse_edges$annotation %in% setbp1_comms,]# filter for only setbp1 communities 
data <- data[data$name %in% setbp1_genes,]# filter for known targets of SETBP1
```

```{r assign-mult-to-ts-setbp1-comm-edges}
mult_genes <- multiplicity[!grepl("-", multiplicity$gene),] # remove genes with alternate alleles aka a "-"
mult_genes <- mult_genes %>% separate(edges, c("tf", "gene")) #grab multiplicity

merged_data <- left_join(data, mult_genes, by = c('name' = 'gene')) # add multiplicity to tse community data

merged_data_gene <- merged_data[merged_data$entity == "gene",] %>% na.omit(.) # filter for only genes

affected_merged <- merged_data_gene[merged_data_gene$tissue_type %in% affected,] # filter for affected tissues
```

```{r visualize-tse-in-setbp1-comm-mult}
# number of tse per tissue
counts <- as.data.frame(table(affected_merged$tissue_type))
counts$Var1 <- str_to_title(counts$Var1) %>% gsub("_", " ", .)

png(
  filename = here("results/tissue_specific_edges/tse_tissues_plot.png"),
  width = 3000,
  height = 2000,
res = 300
)
ggplot(counts, aes(y = Freq, x = Var1)) +
  geom_bar(stat = "identity", fill = "#2C1D6C") + 
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  theme(axis.text.x = element_text(color = "black", angle = 45, vjust = 0.6)) + 
  theme(axis.text.y = element_text(color = "black")) +
  theme(axis.text = element_text(face = "bold")) +
  theme(legend.title = element_text(face = "bold")) +
  theme(legend.text = element_text(face = "bold")) +
  theme(axis.title.y = element_text(face = "bold")) +
  theme(axis.title.x = element_text(face = "bold")) +
  labs(x = "GTEx Affected Tissues", y = "Number of SETBP1's Tissue-specific Edges \n in SETBP1 Communities") +
  theme(title = element_text(face = "bold"),
        plot.title = element_text(hjust = 0, size = 20),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16))
dev.off()

# multiplicity of tissue specific edges of SETBP1 in SEBTP1 communities
counts <- as.data.frame(table(affected_merged$multiplicity))

png(
  filename = here("results/tissue_specific_edges/mult_density_setbp1_tse_plot.png"),
  width = 3000,
  height = 2000,
res = 300
)
ggplot(counts, aes(y = Freq, x = Var1)) +
  geom_bar(stat = "identity", fill = "#2C1D6C") + 
  theme_bw() +
  theme(text = element_text(family = "Helvetica")) +
  theme(axis.text.x = element_text(color = "black", angle = 45, vjust = 0.6)) + 
  theme(axis.text.y = element_text(color = "black")) +
  theme(axis.text = element_text(face = "bold")) +
  theme(legend.title = element_text(face = "bold")) +
  theme(legend.text = element_text(face = "bold")) +
  theme(axis.title.y = element_text(face = "bold")) +
  theme(axis.title.x = element_text(face = "bold")) +
  labs(x = "Multiplicity of SETBP1 Tissue-specific edges \n in SETBP1 Communities", y = "Density") +
  theme(title = element_text(face = "bold"),
        plot.title = element_text(hjust = 0, size = 20),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16))
dev.off()
```


NOTE FOR CR: Need docker version sebtp1_manuscript:1.0.12
Code below for UpSet plot was adapted from vignette [here](https://jokergoo.github.io/ComplexHeatmap-reference/book/upset-plot.html) 
```{r uset-plot-overlapping-edges}
# identifying quartiles for multiplcity
summary(affected_merged$multiplicity)
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#   1.00   25.00   30.00   25.79   31.00   31.00 

# filtering to excude edges that have a multiplicity of less than or equal to Q1
affected_merged <- affected_merged[affected_merged$multiplicity <= 25, ]

# formatting data
subset <- subset(affected_merged, select = c("name", "tissue_type")) # selecting columns of interest

subset$present <- 1# adding a count for sets
subset <- pivot_wider(subset, names_from = name, values_from = present) # pivoting wider
subset <- mutate_all(subset, ~replace_na(.,0)) #change NA to 0
subset$tissue_type <- str_to_title(subset$tissue_type) %>% gsub("_", " ", .)#reformat tissue type
row_names_data <- subset$tissue_type
subset <- subset[,-1]
rownames(subset) <- row_names_data

m1 <- make_comb_mat(subset, mode = "distinct") # distinct mode: 1 means in that set and 0 means not in that set, then 1 1 0
t(m1) #view

# plotting upset plot
png(
  filename = here("results/tissue_specific_edges/upset_plot.png"),
  width = 3000,
  height = 2000,
res = 300
)
UpSet(m1, left_annotation = upset_left_annotation(m1))
dev.off()

# making the combination matrix
subset <- t(subset)

m1 <- make_comb_mat(subset, mode = "distinct") # distinct mode: 1 means in that set and 0 means not in that set, then 1 1 0
t(m1) #view

# plotting upset plot
png(
  filename = here("results/tissue_specific_edges/upset_plot.png"),
  width = 3000,
  height = 2000,
res = 300
)
UpSet(m1, left_annotation = upset_left_annotation(m1))
dev.off()
```

```{r}
# run style
style_file(here("src/tissue_specific_edges/01_tissue_specific_edge_identification.Rmd"))
# lintr was run as well
```

```{r}
fptm <- proc.time() - ptm
fptm[3] / 60
```

```{r}
sessionInfo()
```
