---
title: "alpaca_network_analysis"
output: html_document
date: "2023-06-20"
---

Goal: Identify the overlap between all community and cell type pairings for each tissue as well as only between the Communities containing TF Setbp1. 
* Cerebral cortex: TF Setbp1 is in community 1
* Kidney: TF Setbp1 is in community 1

```{r}
set.seed(2178)
library(netZooR)
library(here)
library(corrplot)
library(dplyr)
library(readr)
library(styler)
library(RColorBrewer)
library(tidyr)
library(stringr)
ptm <- proc.time()
```
# Investigate differences in community structure using Jaccard Index:

## Cortex
```{r}
# Get the list of files in the directory
files <- list.files(here("results","alpaca","main"), pattern = "cortex.*_final_memb.txt")

# Initialize an empty data frame
cortex_memb <- data.frame()

# Loop through the files
for (file in files) {
  
  # Read the data from the file
  data <- read_delim(here("results","alpaca","main", file), col_names = FALSE)  # Modify if your files are not tab-delimited
  
  # Extract the cell type from the filename
  cell_type <- strsplit(file, "_cortex.RData_ALPACA_final_memb.txt")[[1]][1]
  
  # Add a new column to the data for the cell type
  data$cell_type <- cell_type
  
  # Append the data to the master data frame
  cortex_memb <- bind_rows(cortex_memb, data)
}

names(cortex_memb) <- c("node","module","cell_type")
# View the combined data
head(cortex_memb)

```

```{r}
# Assuming 'all_data' is your dataframe with 'node', 'module', and 'cell_type' columns

# Get the unique cell types
unique_cell_types <- unique(cortex_memb$cell_type)

# Initialize a list to store the results
jaccard_results <- list()

# Loop over each unique cell type
for (cell_type in unique_cell_types) {
  # Get the unique modules for this cell type
  unique_modules <- unique(cortex_memb$module[cortex_memb$cell_type == cell_type])
  
  # Loop over each unique module
  for (module in unique_modules) {
    # Get the nodes for the current cell type and module
    current_nodes <- cortex_memb$node[cortex_memb$cell_type == cell_type & cortex_memb$module == module]
    
    # Loop over all other cell types
    for (other_cell_type in unique_cell_types[unique_cell_types != cell_type]) {
      # Get the unique modules for the other cell type
      other_unique_modules <- unique(cortex_memb$module[cortex_memb$cell_type == other_cell_type])
      
      # Loop over each unique module for the other cell types
      for (other_module in other_unique_modules) {
        # Get the nodes for the other cell type and module
        other_nodes <- cortex_memb$node[cortex_memb$cell_type == other_cell_type & cortex_memb$module == other_module]
        
        # Calculate the Jaccard index
        intersection_size <- length(intersect(current_nodes, other_nodes))
        union_size <- length(union(current_nodes, other_nodes))
        jaccard_index <- intersection_size / union_size
        
        # Store the result in the list
        result_key <- paste(cell_type, module, other_cell_type, other_module, sep="_")
        jaccard_results[[result_key]] <- jaccard_index
      }
    }
  }
}

# Convert the results list to a dataframe
jaccard_df <- data.frame(celltype_module = names(jaccard_results), jaccard_index = unlist(jaccard_results), stringsAsFactors = FALSE)

# View the Jaccard results
head(jaccard_df)

setbp1_comm_filtered <- jaccard_df %>%
  # separate the celltype_module column into two new columns: celltype1 and celltype2
  separate(celltype_module, into = c("celltype1", "module1", "celltype2", "module2"), sep = "_")

# filter rows where both module numbers are exactly '1' because Setbp1 the TF is found in community 1 in all cell types for cerebral cortex.
df_filtered <- setbp1_comm_filtered %>%
  filter(module1 == '1' & module2 == '1') %>% select(-module1, -module2)

# create the matrix
matrix_df <- df_filtered %>%
  # reshape the data from long format to wide format
  spread(key = celltype2, value = jaccard_index)

# convert the dataframe to a matrix
jaccard_setbp1_matrix <- as.matrix(matrix_df[-1])
rownames(jaccard_setbp1_matrix) <- matrix_df[[1]]

# Create a PNG file
png(here("results","alpaca","setbp1_cortex_community_corrplot.png"), width = 5000, height = 3000, res = 250)
# Plot the correlogram
my_palette <- colorRampPalette(c("white","#C4DAFD","#7AABE2","#0072B4","#00366C"))
setbp1_cortex_corr <- corrplot(jaccard_setbp1_matrix, method = "circle", is.corr = FALSE, type = 'lower', diag = FALSE, col = my_palette(100))
dev.off()

# Create a PNG file
png(here("results","alpaca","setbp1_cortex_community_histogram.png"), width = 2000, height = 1500, res = 250)
# Plot the histogram
setbp1_cortex_hist <- hist(jaccard_df$jaccard_index, 
                    main = "Distribution of Jaccard Index across celltypes and communities", 
                    xlab = "Jaccard Index", 
                    ylab = "Frequency", 
                    col = "#5F9EFD", 
                    border = "black")
dev.off()

```

## Kidney
```{r}
# Get the list of files in the directory
files <- list.files(here("results","alpaca","main"), pattern = "kidney.*_final_memb.txt")

# Initialize an empty data frame
kidney_memb <- data.frame()

# Loop through the files
for (file in files) {
  
  # Read the data from the file
  data <- read_delim(here("results","alpaca","main", file), col_names = FALSE)  # Modify if your files are not tab-delimited
  
  # Extract the cell type from the filename
  cell_type <- strsplit(file, "_kidney.RData_ALPACA_final_memb.txt")[[1]][1]
  
  # Add a new column to the data for the cell type
  data$cell_type <- cell_type
  
  # Append the data to the master data frame
  kidney_memb <- bind_rows(kidney_memb, data)
}

names(kidney_memb) <- c("node","module","cell_type")
# View the combined data
head(kidney_memb)

```

```{r}
# Assuming 'all_data' is your dataframe with 'node', 'module', and 'cell_type' columns

# Get the unique cell types
unique_cell_types <- unique(kidney_memb$cell_type)

# Initialize a list to store the results
jaccard_results <- list()

# Loop over each unique cell type
for (cell_type in unique_cell_types) {
  # Get the unique modules for this cell type
  unique_modules <- unique(kidney_memb$module[kidney_memb$cell_type == cell_type])
  
  # Loop over each unique module
  for (module in unique_modules) {
    # Get the nodes for the current cell type and module
    current_nodes <- kidney_memb$node[kidney_memb$cell_type == cell_type & kidney_memb$module == module]
    
    # Loop over all other cell types
    for (other_cell_type in unique_cell_types[unique_cell_types != cell_type]) {
      # Get the unique modules for the other cell type
      other_unique_modules <- unique(kidney_memb$module[kidney_memb$cell_type == other_cell_type])
      
      # Loop over each unique module for the other cell types
      for (other_module in other_unique_modules) {
        # Get the nodes for the other cell type and module
        other_nodes <- kidney_memb$node[kidney_memb$cell_type == other_cell_type & kidney_memb$module == other_module]
        
        # Calculate the Jaccard index
        intersection_size <- length(intersect(current_nodes, other_nodes))
        union_size <- length(union(current_nodes, other_nodes))
        jaccard_index <- intersection_size / union_size
        
        # Store the result in the list
        result_key <- paste(cell_type, module, other_cell_type, other_module, sep="_")
        jaccard_results[[result_key]] <- jaccard_index
      }
    }
  }
}

# Convert the results list to a dataframe
jaccard_df <- data.frame(celltype_module = names(jaccard_results), jaccard_index = unlist(jaccard_results), stringsAsFactors = FALSE)

# View the Jaccard results
head(jaccard_df)

setbp1_comm_filtered <- jaccard_df %>%
  # separate the celltype_module column into two new columns: celltype1 and celltype2
  separate(celltype_module, into = c("celltype1", "module1", "celltype2", "module2"), sep = "_")

# filter rows where both module numbers are exactly '1' because Setbp1 the TF is found in community 1 in all cell types for kidney.
df_filtered <- setbp1_comm_filtered %>%
  filter(module1 == '1' & module2 == '1') %>% select(-module1, -module2)

# create the matrix
matrix_df <- df_filtered %>%
  # reshape the data from long format to wide format
  spread(key = celltype2, value = jaccard_index)

# convert the dataframe to a matrix
jaccard_setbp1_matrix <- as.matrix(matrix_df[-1])
rownames(jaccard_setbp1_matrix) <- matrix_df[[1]]

# Create a PNG file
png(here("results","alpaca","setbp1_kidney_community_corrplot.png"), width = 5000, height = 3000, res = 250)
# Plot the correlogram
setbp1_kidney_corr <- corrplot(jaccard_setbp1_matrix, method = "circle", is.corr = FALSE, type = 'lower', diag = FALSE)
dev.off()

# Create a PNG file
png(here("results","alpaca","setbp1_kidney_community_histogram.png"), width = 2000, height = 1500, res = 250)
# Plot the histogram
setbp1_kidney_hist <- hist(jaccard_df$jaccard_index, 
                    main = "Distribution of Jaccard Index across celltypes and communities", 
                    xlab = "Jaccard Index", 
                    ylab = "Frequency", 
                    col = "#FFB000", 
                    border = "black")
dev.off()
```

```{r processing-time}
fptm <- proc.time() - ptm
fptm[3] / 60
```
 elapsed 
2.055817 

```{r style-file}
# run style
style_file(here("src/network_scripts/decoupleR/03_TF_activity_kidney.Rmd")) # commented out after being run once
# lintr was run as well
```

```{r session-info}
sessionInfo()
```
R version 4.1.3 (2022-03-10)
Platform: x86_64-pc-linux-gnu (64-bit)
Running under: Ubuntu 20.04.6 LTS

Matrix products: default
BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/openblas-pthread/libopenblasp-r0.3.8.so

locale:
[1] C

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] stringr_1.5.0       tidyr_1.3.0         styler_1.9.0        readr_2.1.4         dplyr_1.1.0         corrplot_0.92       here_1.0.1          netZooR_1.2.1      
 [9] matrixcalc_1.0-6    yarn_1.20.0         pandaR_1.26.0       Biobase_2.54.0      BiocGenerics_0.40.0 reticulate_1.28     igraph_1.4.1       

loaded via a namespace (and not attached):
  [1] utf8_1.2.3                  R.utils_2.12.2              proto_1.0.0                 RUnit_0.4.32                tidyselect_1.2.0           
  [6] RSQLite_2.3.0               AnnotationDbi_1.56.2        grid_4.1.3                  BiocParallel_1.28.3         munsell_0.5.0              
 [11] base64url_1.4               codetools_0.2-18            preprocessCore_1.56.0       chron_2.3-59                pbdZMQ_0.3-9               
 [16] withr_2.5.0                 colorspace_2.1-0            Category_2.60.0             filelock_1.0.2              knitr_1.42                 
 [21] uuid_1.1-0                  rstudioapi_0.14             stats4_4.1.3                MatrixGenerics_1.6.0        repr_1.1.6                 
 [26] GenomeInfoDbData_1.2.7      bit64_4.0.5                 rhdf5_2.38.1                rprojroot_2.0.3             downloader_0.4             
 [31] vctrs_0.5.2                 generics_0.1.3              xfun_0.37                   BiocFileCache_2.2.1         R6_2.5.1                   
 [36] doParallel_1.0.17           GenomeInfoDb_1.30.1         illuminaio_0.36.0           RJSONIO_1.3-1.8             locfit_1.5-9.7             
 [41] bitops_1.0-7                rhdf5filters_1.6.0          cachem_1.0.7                reshape_0.8.9               DelayedArray_0.20.0        
 [46] assertthat_0.2.1            vroom_1.6.1                 BiocIO_1.4.0                scales_1.2.1                nnet_7.3-17                
 [51] gtable_0.3.1                rlang_1.0.6                 genefilter_1.76.0           splines_4.1.3               rtracklayer_1.54.0         
 [56] GEOquery_2.62.2             hexbin_1.28.2               reshape2_1.4.4              yaml_2.3.7                  GenomicFeatures_1.46.5     
 [61] backports_1.4.1             RBGL_1.70.0                 tools_4.1.3                 nor1mix_1.3-0               ggplot2_3.4.1              
 [66] ellipsis_0.3.2              gplots_3.1.3                RColorBrewer_1.1-3          ggdendro_0.1.23             siggenes_1.68.0            
 [71] STRINGdb_2.6.5              gsubfn_0.7                  Rcpp_1.0.10                 hash_2.2.6.2                plyr_1.8.8                 
 [76] base64enc_0.1-3             sparseMatrixStats_1.6.0     progress_1.2.2              zlibbioc_1.40.0             purrr_1.0.1                
 [81] RCurl_1.98-1.10             prettyunits_1.1.1           openssl_2.0.5               sqldf_0.4-11                bumphunter_1.36.0          
 [86] S4Vectors_0.32.4            cluster_2.1.2               SummarizedExperiment_1.24.0 fs_1.6.1                    magrittr_2.0.3             
 [91] data.table_1.14.8           R.cache_0.16.0              matrixStats_0.63.0          hms_1.1.2                   evaluate_0.20              
 [96] xtable_1.8-4                XML_3.99-0.13               mclust_6.0.0                IRanges_2.28.0              compiler_4.1.3             
[101] biomaRt_2.50.3              minfi_1.40.0                tibble_3.1.8                RCy3_2.14.2                 KernSmooth_2.23-20         
[106] crayon_1.5.2                R.oo_1.25.0                 htmltools_0.5.4             GOstats_2.60.0              mgcv_1.8-39                
[111] tzdb_0.3.0                  DBI_1.1.3                   dbplyr_2.3.1                MASS_7.3-55                 rappdirs_0.3.3             
[116] Matrix_1.5-3                permute_0.9-7               cli_3.6.0                   quantro_1.28.0              uchardet_1.1.1             
[121] R.methodsS3_1.8.2           quadprog_1.5-8              parallel_4.1.3              GenomicRanges_1.46.1        pkgconfig_2.0.3            
[126] GenomicAlignments_1.30.0    IRdisplay_1.1               xml2_1.3.3                  foreach_1.5.2               annotate_1.72.0            
[131] penalized_0.9-52            rngtools_1.5.2              multtest_2.50.0             beanplot_1.3.1              XVector_0.34.0             
[136] AnnotationForge_1.36.0      doRNG_1.8.6                 scrime_1.3.5                digest_0.6.31               vegan_2.6-4                
[141] graph_1.72.0                Biostrings_2.62.0           base64_2.0.1                edgeR_3.36.0                DelayedMatrixStats_1.16.0  
[146] GSEABase_1.56.0             restfulr_0.0.15             curl_5.0.0                  Rsamtools_2.10.0            gtools_3.9.4               
[151] rjson_0.2.21                lifecycle_1.0.3             nlme_3.1-155                jsonlite_1.8.4              Rhdf5lib_1.16.0            
[156] viridisLite_0.4.1           askpass_1.1                 limma_3.50.3                fansi_1.0.4                 pillar_1.8.1               
[161] lattice_0.20-45             plotrix_3.8-2               KEGGREST_1.34.0             fastmap_1.1.1               httr_1.4.5                 
[166] survival_3.3-1              GO.db_3.14.0                glue_1.6.2                  png_0.1-8                   iterators_1.0.14           
[171] bit_4.0.5                   Rgraphviz_2.38.0            stringi_1.7.12              HDF5Array_1.22.1            blob_1.2.3                 
[176] org.Hs.eg.db_3.14.0         caTools_1.18.2              memoise_2.0.1               IRkernel_1.3.2             
