---
title: "README"
author: "Jordan Whitlock"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Network Construction and Analysis

## Purpose:
All scripts in this directory are used to build necessary inputs required for network construction as well as downstream analyses.

### Reproducibility:

**Network Construction**
* PANDA jobs were run using UAB Cheaha Supercomputer and it's SLURM scheduler. Jobs were run as an array, within a [Docker](https://hub.docker.com/repository/docker/jordanwhitlock/jw_diseasenetworks/general) container converted Singularity in order to execute.
* Human Protein-Protein Interaction (PPI) and Transcription Factor Motif (TF-Motif) inputs were previously generated in [Whitlock et al](). The github repository for constructing both inputs and associated code is found [here](https://github.com/lasseignelab/230227_JW_Setbp1Manuscript/tree/main) for both the [Human TF-Motif](https://github.com/lasseignelab/230227_JW_Setbp1Manuscript/blob/main/src/network_scripts/PANDA_input_construction/02_Human_TFmotif.Rmd) and [Human PPI](https://github.com/lasseignelab/230227_JW_Setbp1Manuscript/blob/main/src/network_scripts/PANDA_input_construction/04_HumanMouse_ppi.Rmd)
* Expression data was obtained from GTEx using Recount3 (accessed: 220806-220807) and includes all tissue types except blood and study_na

### Scripts:

#### PANDA Expression Inputs and Network Construction:
This directory contains all scripts needed to construct [PANDA](https://netzoo.github.io/zooanimals/panda/) regulatory networks. In addition, the *.err* and *.out* files for each array job are included here to provide detailed information on the jobs. Prior to building PANDA networks, the user must first construct the expression inputs needed and then build the *.txt*, which is a list of expression input file paths needed for the array job.

```{r echo = FALSE}
fs::dir_tree("GTEx_PANDA")
```